### 2019-03-31 conclusion.tex
Randoop case study

\section{Conclusion}\label{sec:conclusion}

\todo{TODO}

\todo{\TheDeterminismChecker found errors that the
  Randoop developers had missed, despite spending over 100 hours on the
  problem of nondeterminism.}