### 2019-04-06 conclusion.tex

\section{Conclusion}\label{sec:conclusion}

We designed a type system that helps programmers express determinism specifications
in sequential programs. 
We implemented our type system in Java, applied it to real world 
software and found it to be effective in finding previously unknown determinism bugs. 
\TheDeterminismChecker found errors that the
Randoop developers had missed, despite spending over 100 hours on the
problem of nondeterminism. We compared our tool with existing state of the art flake test detectors
and found that \theDeterminismChecker found all the bugs arising due to nondeterminism
in sequential programs that were used for evaluation in NonDex~\cite{nondex} and DeFlaker~\cite{deflaker}.
To the best of our knowledge, ours is the
first compile time verification approach addressing the problem of nondeterminism.